Stochastic blockmodeling of relational event dynamics

### 2012-06-01 docs/discussion.tex
Tweaked the discussion section.

%PS: I would prefer to say: In this paper, we propose (or we proposed)….
We propose a hierarchical approach for modeling event-based network data as a continuous-time Markov process. 
Our approach generalizes traditional static notions of stochastic equivalence on nodes (such as stochastic blockmodels) to a dynamic formulation. 
The proposed model posits the existence of groups of nodes, such that all members of a group are governed by the same dynamic process, and groups are differentiated by having different dynamic processes. 
%PS: sentence below is incomplete (after the first "the")…..I decided to take it out anyway..
% Given a specification for the, the proposed method employs latent variables to model the heterogeneity in the event dynamics.
%is analogous to recent hierarchical extensions for latent position models \cite{Handcock2007} and exponential random graph models \cite{Schweinberger2011}.
%The method combines a latent variable framework (stochastic blockmodels) with a local dependence model for event sequences (relational event models \cite{Butts2008}).
%This provides detailed models of event dynamics among subsets of nodes while allowing for heterogeneity in the dynamics of the network as a whole.

%This analysis suggests 
Our approach has the ability to uncover systematic differences in dynamic behaviors among subsets of nodes.
% across a variety of data sets.
In Section \ref{sec:experiments} we show the model has improved predictive accuracy over baseline methods on real data with respect to ranking tasks and the likelihood of unobserved data.
In particular, we show that our proposed approach leads to improved predictive accuracy when compared to both (a) relational event models that lack latent clusters, and (b) stochastic blockmodels for count data, both of which are both special cases of our model.
Though prediction is not the main focus of the model, the results provide some evidence that having latent structure (i.e. K>1) can lead to improvements in predictive power. Our prediction results show that it is easy to overfit with this model. 
%If we restrict to K=2 on these small data sets, we systematically outperform the K=1 model. 
%{\red (TODO: confirm this)}
 As K increases beyond 2 (e.g., for the unrestrained CRP prior), the model is prone to overfit. This is not surprising, given that the number of fitted parameters scales as the square of the number of components K in the model. A natural future direction worth exploring for this model is a class of priors that are more resistant to overfitting.

Another direction to explore is that of including different statistics in the specification of $\mathbf{s}(t,i,j,\mathcal{A}_t)$, and using our approach to study how the roles of these statistics vary across nodes.  
One potentially interesting extension, analogous to \cite{Airoldi2008}, would be to allow the latent class $z_i$ to be drawn from node-specific membership vectors $\pi_i$  after each change point---we leave this to future work.

>>>>

%PS: I would prefer to say: In this paper, we propose (or we proposed)….
%We propose a hierarchical approach for modeling event-based network data as a continuous-time Markov process. 
We have here introduced a family of relational event models that can flexibly capture heterogeneity in the underlying interaction dynamics.
Our approach generalizes traditional, static notions of stochastic equivalence on nodes (such as stochastic blockmodels) to the dynamic context. 
The proposed model family posits the existence of groups of nodes, such that all members of a group are governed by the same dynamic process, and groups are differentiated by having different dynamic processes. 
%PS: sentence below is incomplete (after the first "the")…..I decided to take it out anyway..
% Given a specification for the, the proposed method employs latent variables to model the heterogeneity in the event dynamics.
%is analogous to recent hierarchical extensions for latent position models \cite{Handcock2007} and exponential random graph models \cite{Schweinberger2011}.
%The method combines a latent variable framework (stochastic blockmodels) with a local dependence model for event sequences (relational event models \cite{Butts2008}).
%This provides detailed models of event dynamics among subsets of nodes while allowing for heterogeneity in the dynamics of the network as a whole.

%This analysis suggests 
Our approach has the ability to uncover systematic differences in dynamic behaviors among subsets of nodes, even in the absence of differences in marginal interaction rates.
% across a variety of data sets.
The analyses of Section \ref{sec:experiments} show that this model family has improved predictive accuracy over baseline methods on real data with respect to ranking tasks and the likelihood of unobserved data.
In particular, our proposed approach leads to improved predictive accuracy when compared to both (a) relational event models that lack latent clusters, and (b) stochastic blockmodels for count data, both of which are special cases of our model family.
Though prediction is not the main focus of our effort, these results provide evidence that having latent structure (i.e. K>1) can lead to improvements in predictive power. However, our prediction results also show that it is easy to overfit with this model family. 
%If we restrict to K=2 on these small data sets, we systematically outperform the K=1 model. 
%{\red (TODO: confirm this)}
 As K increases beyond 2 (e.g., for the unrestrained CRP prior), these models are prone to overfit. This is not surprising, given that the number of fitted parameters scales as the square of the number of components K in the model. A natural direction worth exploring for this family is a class of priors that are more resistant to overfitting (e.g., by imposing more structure on the parameters within each group).

Another direction to explore is that of including different statistics in the specification of $\mathbf{s}(t,i,j,\mathcal{A}_t)$, and using our approach to study how the roles of these statistics vary across nodes.  
One potentially interesting extension, analogous to \cite{Airoldi2008}, would be to allow the latent class $z_i$ to be drawn from node-specific membership vectors $\pi_i$  after each change point.  Allowing underlying dynamics to be governed by a latent membership structure opens the door to a wide range of possibilities, with many opportunities for further elaboration.
